### 2022-04-05 MSA Project.Rmd
Initial commit

---
title: "MSA Project"
author: "Niko Miller"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Lorem Ipsum