### 2022-04-10 MSA-Project.Rmd
Updated variable description table and added references for tables and figures

---
title: "PCA on Finnish Municipalities"
author: "Niko Miller"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  bookdown::pdf_document2:
    keep_tex: false
    number_sections: true
    fig_caption: true
    citation_package: natbib
bibliography: references.bib
header-includes:
- \pagenumbering{gobble}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{mathtools}
- \usepackage{graphicx}
- \usepackage{placeins}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{framed}
- \usepackage[font={small}]{caption}
- \usepackage[font={small,it}]{subcaption}
- \usepackage{hyperref}
- \usepackage[labelsep=period]{caption}
- \usepackage[labelfont=bf]{caption}
- \captionsetup{justification=raggedright,singlelinecheck=false}

subtitle: "Dimensionality reduction with PCA on Finnish municipality demographics"
geometry: top=0.7in, bottom=0.5in, left=0.5in, right=0.7in
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
options(ggrepel.max.overlaps = Inf)

library(tidyverse)
library(knitr)
library(kableExtra)
library(summarytools)
library(pxweb)
library(ggplot2)
library(reshape2)
library(ggcorrplot)
library(ggrepel)
library(ggfortify)

# muni.latest <- read.csv("muni.csv")
# muni.latest <- muni.latest[, -1] %>% as_tibble()

muni.latest <- read.csv("muni_2021.csv", sep = ";" , header = F)
colnames(muni.latest) <- c('Area',
                           'Deg.Urbanisation',
                           'Popul',
                           'Popul.Growth',
                           'Prop.Under15',
                           'Prop.15to64',
                           'Prop.Over64',
                           'Prop.Swedish',
                           'Prop.Foreign',
                           'Excess.Births',
                           'Migr.Gain',
                           'Families',
                           'Households',
                           'Prop.Households.Terr.Det',
                           'Prop.Households.Rental',
                           'Prop.Educ.Degree2',
                           'Prop.Educ.Degree3',
                           'Labour.Force',
                           'Empl.Rate',
                           'Prop.Empl.Muni',
                           'Prop.Unempl',
                           'Prop.Pensioners',
                           'Depend.Ratio',
                           'Jobs.Muni',
                           'Prop.Primary.Sector',
                           'Prop.Secondary.Sector',
                           'Prop.Services.Sector',
                           'Jobs.Self.Suff',
                           'Contr.Margin',
                           'Loan.Stock',
                           'Group.Loan.Stock',
                           'Educ.Cult.Activity',
                           'Soc.Health.Activity')

muni.latest <- muni.latest %>% filter(!grepl('koko maa|maakunta|seutukunta|mariehamn', Area, ignore.case = T))

muni.latest <- muni.latest %>% 
  arrange(desc(Popul))

```

<!-- \vspace{-1cm} -->

<!-- \tableofcontents -->

<!-- \listoftables -->
<!-- \listoffigures -->

\newpage
\pagenumbering{arabic}

# Introduction

## Data and Research Questions

Tilastokeskus provides data on 32 demographic variables of Finnish municipalities. The data can be freely obtained from Tilastokeskus' website in csv. format [@data].

The variables provide information on e.g., the municipalities' population and its growth, age structure, economic sector structure, and the split between different housing types.

The purpose of this study is to reduce dimensionality in the data and assess whether differences between the municipalities can be explained by considerably less than 32 dimensions - namely by first few principal components (PCs). This study's contribution can be seen as exploratory data analysis where the municipalities are clustered in a lower dimension. The results can be applied in further work in e.g., more advanced clustering, prediction or economic decision making.

Clustering municipalities with demographic factors can be useful in a business context, e.g., marketing. Knowing how municipalities differ in demographics can help a company to improve its targeting and e.g., launch different marketing campaigns or product lines in different municipalities.

# Univariate Analysis

## Variable Descriptions

Table \@ref(tab:vardesc) describes the variables used in this study.

```{r vardesc, echo=FALSE}

vars.orig <- c('Degree of urbanisation',
               'Population',
               'Population change from the previous year',
               'Share of persons aged under 15 of the population',
               'Share of persons aged 15 to 64 of the population',
               'Share of persons aged over 64 of the population',
               'Share of Swedish-speakers of the population',
               'Share of foreign citizens of the population',
               'Excess of births',
               'Intermunicipial migration gain/loss',
               'Number of families',
               'Number of household-dwelling units',
               'Share of household-dwelling units living in terraced houses and detached houses',
               'Share of household-dwelling units living in rental dwellings',
               'Share of persons aged 15 or over with at least upper secondary qualifications',
               'Share of persons aged 15 or over with tertiary level qualifications',
               'Employed labour force resident in the area',
               'Employment rate',
               'Share of persons working in their municipality of residence',
               'Proportion of unemployed among the labour force',
               'Proportion of pensioners of the population',
               'Economic dependency ratio',
               'Number of workplaces in the area',
               'Share of workplaces in primary production',
               'Share of workplaces in secondary production',
               'Share of workplaces in services',
               'Workplace self-sufficiency',
               'Annual contribution margin',
               'Loan stock',
               'Group loan stock',
               'Educational and cultural activities',
               'Social and health care activities')

vars <-  c('Deg.Urbanisation',
           'Popul',
           'Popul.Growth',
           'Prop.Under15',
           'Prop.15to64',
           'Prop.Over64',
           'Prop.Swedish',
           'Prop.Foreign',
           'Excess.Births',
           'Migr.Gain',
           'Families',
           'Households',
           'Prop.Households.Terr.Det',
           'Prop.Households.Rental',
           'Prop.Educ.Degree2',
           'Prop.Educ.Degree3',
           'Labour.Force',
           'Empl.Rate',
           'Prop.Empl.Muni',
           'Prop.Unempl',
           'Prop.Pensioners',
           'Depend.Ratio',
           'Jobs.Muni',
           'Prop.Primary.Sector',
           'Prop.Secondary.Sector',
           'Prop.Services.Sector',
           'Jobs.Self.Suff',
           'Contr.Margin',
           'Loan.Stock',
           'Group.Loan.Stock',
           'Educ.Cult.Activity',
           'Soc.Health.Activity')

years <- c(2020,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2019,
           2020,
           2020,
           2020,
           2019,
           2019,
           2019,
           2019,
           2019,
           2020,
           2020,
           2020,
           2020,
           2020)

units <- c("%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "%",
           "%",
           "Persons",
           "Persons",
           "Number",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Ratio",
           "Number",
           "%",
           "%",
           "%",
           "Ratio",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita")

var.df <- data.frame(vars, vars.orig, years, units)

kable(var.df, format = "latex", align = "lll", col.names = c("Variable", "Original Variable","Year","Unit"), caption = "Description of all used variables. Variable is the name of the variable that is used in this report. Original Variable is the original variable given in Tilastokeskus' dataset, Year is the timestamp of the data point, and Unit is the unit in which the variable is measured in.") %>%
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = c("HOLD_position", "scale_down"), position = "left")

```

## Descriptive Statistics

Table \@ref(tab:descstats) shows descriptive statistics for each variable.

```{r descstats, echo=FALSE}
stats <- muni.latest[, -1] %>% apply(2, summary) %>% t() %>% as.data.frame()
stats$Std <- muni.latest[, -1] %>% apply(2, sd)

stats.df <- data.frame(vars, stats)
rownames(stats.df) <- NULL

kable(stats.df, format = "latex", align = "lrrrrrr", caption = "Descriptive Statistics of the variables. Variable is the name of the variable, Min is the minimum, 1st Qu. is the 1st quartile, Median is the median, Mean is the arithmetic mean, 3rd Qu. is the 3rd quartile, Max is the maximum and Std. is the sample standard deviation.", digits = 1, col.names = c("Variable", "Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max", "Std.")) %>% 
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position", position = "left")

```

## Distribution Plots

```{r, include=FALSE}

ages <- muni.latest %>% 
  select(Prop.Under15, Prop.15to64, Prop.Over64)
ages <- melt(ages)

living <- muni.latest %>% 
  select(Prop.Households.Terr.Det, Prop.Households.Rental)
living <- melt(living)

employment <- muni.latest %>% 
  select(Empl.Rate, Prop.Unempl, Prop.Pensioners)
employment <- melt(employment)

sectors <- muni.latest %>% 
  select(Prop.Primary.Sector, Prop.Secondary.Sector, Prop.Services.Sector)
sectors <- melt(sectors)

```

To analyze how variables are distributed, we use Kernel density estimation to estimate the probability densities for several variable groups.

Figure \@ref(fig:agekde) shows the probability density estimates for age groups. Age groups are shares of population aged below 15, 15 to 64, and over 64. We can see that ...

Figure \@ref(fig:housingkde) shows the probability density estimates for housing types. Housing types are shares of population living in either terraced or detached houses and in rental apartments. The Figure revels that ...

Figure \@ref(fig:employmentkde) shows the probability density estimates for employment measures. Employment measures are the employment rate, the proportion of population that are unemployed, and the proportion of population that are pensioners. The distributions show that ...

Figure \@ref(fig:sectorkde) shows the probability density estimates for the weights of economic sectors in the municipalities. Sectors are the primary-, secondary-, and services sectors. We can interpret that ...

```{r agekde, echo=FALSE, fig.cap='Probability density estimates for age groups', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r housingkde, echo=FALSE, fig.cap='Probability density estimates for housing types', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Housing Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r employmentkde, echo=FALSE, fig.cap='Probability density estimates for employment related measures',fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Measures") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r sectorkde, echo=FALSE, fig.cap='Probability density estimates for economic sectors', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(sectors, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

# Bivariate Analysis

## Linear Dependencies

To analyze bivariate dependencies in the data, we assess linear relationships between the variables with Pearson's correlation analysis. Figure \@ref(fig:corrplot) shows a heatmap of correlations between all variables.

The heatmap shows that ...

```{r corrplot, echo=FALSE, fig.dim=c(18,18), fig.cap="Heatmap of Pearson'n correlation between all variables", optipng = '-o7'}
corr <- cor(muni.latest[, -1])
ggcorrplot(corr, hc.order = TRUE, type = "full", lab = TRUE, title = "Heatmap of Correlations", lab_size = 3.5)

```

# Principal Component Analysis (PCA)

To assess how much proportion of variance can be explained with the first few PCs, we create a scree plot. Figure \@ref(fig:screeplot) shows the resulting plot. We can see that we can explain close to 55% of all variation using only the first two PCs. Furthermore, close to 70% of the variation can be explained using the first four PCs.

## Scree Plot

```{r screeplot, echo=FALSE, fig.cap="Variance explained by each Princial Component (PC) up to the 20th PC. This type of a plot is also known as a Scree Plot", optipng = '-o7'}

muni.latest.pca <- muni.latest %>% column_to_rownames(var = "Area")
muni.pca <- prcomp(muni.latest.pca, center = T, scale. = T)

muni.pca.summary <- summary(muni.pca)
scree.df <- muni.pca.summary$importance[, 1:20] %>% t() %>% as.data.frame()
scree.df$PC <- 1:20
scree.df$Included <- c(rep("Yes",4), rep("No",16))
scree.df <- scree.df[, c(4,2,3,5)]
rownames(scree.df) <- NULL
colnames(scree.df) <- c("PC", "Prop.var", "Cum.prop", "Included")

ggplot(scree.df) +
  geom_bar(aes(x = PC, y = Prop.var, fill = Included), stat = "identity") +
  scale_fill_manual(values = c("#440154", "#3dbc74")) +
  geom_line(aes(x = PC, y = Cum.prop)) +
  geom_point(aes(x = PC, y = Cum.prop)) +
  geom_hline(yintercept = 0.75, linetype = 2) +
  annotate("text", x=10.5, y=0.78, label= "75% of variance explained", size = 3) +
  labs(x = "Principal Component", y = "Variance Explained (%)", title = "PCA Scree Plot") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_light()

```

## Principal Components 1 and 2

Figure \@ref(fig:scoreplot1) shows the scores for PCs 1 and 2.

Figure \@ref(fig:loadingsplot1) shows the loading directions for PCs 1 and 2.

Figure \@ref(fig:scoreplot2) shows the scores for PCs 3 and 4.

Figure \@ref(fig:loadingsplot2) shows the loading directions for PCs 3 and 4.
```{r scoreplot1, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC1 and PC2", fig.align='center'}

loads <- muni.pca$x[, 1:2] %>% as.data.frame() %>% rownames_to_column() %>% as_tibble() %>% dplyr::rename(Area = rowname)

greatest5.PC1 <- loads %>% 
  arrange(desc(PC1)) %>% 
  head(5)

smallest5.PC1 <- loads %>% 
  arrange(PC1) %>% 
  head(5)

greatest5.PC2 <- loads %>% 
  arrange(desc(PC2)) %>% 
  head(5)

smallest5.PC2 <- loads %>% 
  arrange(PC2) %>% 
  head(5)

loads.reduced <- full_join(greatest5.PC1, smallest5.PC1) %>% full_join(greatest5.PC2) %>% full_join(smallest5.PC2)

ggplot(loads.reduced, aes(x = PC1, y = PC2, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-25,10)) +
  ylim(c(-20,15)) +
  ggtitle("PCA Scores") +
  theme_light()

```

```{r loadingsplot1, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC1 and PC2", fig.align='center'}

ggplot2::autoplot(muni.pca, 
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.4, 0.4)) +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  labs(x = "Standardized PC1 (33.6% explained var.)",
       y = "Standardized PC1 (20.9% explained var.)",
       title = "PCA Loadings") +
  theme_light()

```

## Principal Components 3 and 4

```{r scoreplot2, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC3 and PC4", fig.align='center'}

loads <- muni.pca$x[, 3:4] %>% as.data.frame() %>% rownames_to_column() %>% as_tibble() %>% dplyr::rename(Area = rowname)

greatest5.PC3 <- loads %>% 
  arrange(desc(PC3)) %>% 
  head(5)

smallest5.PC3 <- loads %>% 
  arrange(PC3) %>% 
  head(5)

greatest5.PC4 <- loads %>% 
  arrange(desc(PC4)) %>% 
  head(5)

smallest5.PC4 <- loads %>% 
  arrange(PC4) %>% 
  head(5)

loads.reduced <- full_join(greatest5.PC3, smallest5.PC3) %>% full_join(greatest5.PC4) %>% full_join(smallest5.PC4)

ggplot(loads.reduced, aes(x = PC3, y = PC4, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-5,5)) +
  ylim(c(-8,8)) +
  ggtitle("PCA Scores") +
  theme_light()

```

```{r loadingsplot2, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC3 and PC4", fig.align='center'}

ggplot2::autoplot(muni.pca,
                  x = 3,
                  y = 4,
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.3, 0.3)) +
  scale_y_continuous(limits = c(-0.3, 0.3)) +
  labs(x = "Standardized PC3 (10.2% explained var.)",
       y = "Standardized PC4 (7.2% explained var.)",
       title = "PCA Loadings") +
  theme_light()

```

# Discussion and Conclusions

- PC1 and PC2 do not explain a high percentage of the variance (above 80%). Hence, PCA might not be so well suited
- Using population, employed persons and some other absolute measures might have alienated Helsinki from the other cities and biased the analysis a bit. Perhaps leaving out Helsinki would have been a wise choice
- Highly correlated variables could have been reduced to only one, i.e., remove unnecessary variables
- Something else?

\newpage
# References