### 2022-04-10 MSA-Project.Rmd
Updated report

---
title: "PCA on Finnish Municipalities"
author: "Niko Miller"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  bookdown::pdf_document2:
    keep_tex: false
    number_sections: true
    fig_caption: true
    citation_package: natbib
bibliography: references.bib
header-includes:
- \pagenumbering{gobble}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{mathtools}
- \usepackage{graphicx}
- \usepackage{placeins}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{framed}
- \usepackage[font={small}]{caption}
- \usepackage[font={small,it}]{subcaption}
- \usepackage{hyperref}
- \usepackage[labelsep=period]{caption}
- \usepackage[labelfont=bf]{caption}
- \captionsetup{justification=raggedright,singlelinecheck=false}

subtitle: "Dimensionality reduction on demographic data of Finnish municipalities with PCA"
geometry: top=0.7in, bottom=0.5in, left=0.5in, right=0.7in
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
options(ggrepel.max.overlaps = Inf)

library(tidyverse)
library(knitr)
library(kableExtra)
library(summarytools)
library(pxweb)
library(ggplot2)
library(reshape2)
library(ggcorrplot)
library(ggrepel)
library(ggfortify)

# muni.latest <- read.csv("muni.csv")
# muni.latest <- muni.latest[, -1] %>% as_tibble()

muni.latest <- read.csv("muni_2021.csv", sep = ";" , header = F)
colnames(muni.latest) <- c('Area',
                           'Deg.Urbanisation',
                           'Popul',
                           'Popul.Growth',
                           'Prop.Under15',
                           'Prop.15to64',
                           'Prop.Over64',
                           'Prop.Swedish',
                           'Prop.Foreign',
                           'Excess.Births',
                           'Migr.Gain',
                           'Families',
                           'Households',
                           'Prop.Households.Terr.Det',
                           'Prop.Households.Rental',
                           'Prop.Educ.Degree2',
                           'Prop.Educ.Degree3',
                           'Labour.Force',
                           'Empl.Rate',
                           'Prop.Empl.Muni',
                           'Prop.Unempl',
                           'Prop.Pensioners',
                           'Depend.Ratio',
                           'Jobs.Muni',
                           'Prop.Primary.Sector',
                           'Prop.Secondary.Sector',
                           'Prop.Services.Sector',
                           'Jobs.Self.Suff',
                           'Contr.Margin',
                           'Loan.Stock',
                           'Group.Loan.Stock',
                           'Educ.Cult.Activity',
                           'Soc.Health.Activity')

muni.latest <- muni.latest %>% filter(!grepl('koko maa|maakunta|seutukunta|mariehamn', Area, ignore.case = T))

muni.latest <- muni.latest %>% 
  arrange(desc(Popul))

```

<!-- \vspace{-1cm} -->

<!-- \tableofcontents -->

<!-- \listoftables -->
<!-- \listoffigures -->

\newpage
\pagenumbering{arabic}

# Introduction

## Data Collection and Cleaning

Tilastokeskus provides data on 32 demographic variables of Finnish municipalities. The data can be freely obtained from Tilastokeskus' website in csv. format [@data].

The variables provide information on e.g., the municipalities' population and its growth, age structure, economic sector structure, and the split between different housing types.

We concentrate on municipalities only, so we filtered out the country (koko maa), all counties (maakunta) and subregions (seutukunta) from the data. Moreover, we removed Mariehamn Stad as it contains the same information as Maarianhamina.

## Research Question and Motivation

The purpose of this study is to reduce dimensionality in the data and assess whether differences between the municipalities can be explained by considerably less than 32 dimensions - namely by first few principal components (PCs). This study's contribution can be seen as exploratory data analysis where the municipalities are clustered in a lower dimension. The results can be applied in further work in e.g., more advanced clustering, prediction or economic decision making.

Clustering municipalities with demographic factors can be useful in a business context, e.g., marketing. Knowing how municipalities differ in demographics can help a company to improve its targeting and e.g., launch different marketing campaigns or product lines in different municipalities.

# Univariate Analysis

## Variable Descriptions

Table \@ref(tab:vardesc) describes the variables used in this study. We can see that most of the variables are measured in 2020 and represent a proportion of a municipality's population. Some variables, e.g., Population, have newer data (2021) while a some data points to 2019 (e.g., split of economic sectors). Overall, we have a quite wide demographic spectrum as the variables cover many aspects of the municipalities' nature: population and migration trends, education level and housing, employment, ethnicity, economy, etc.

```{r vardesc, echo=FALSE}

vars.orig <- c('Degree of urbanisation',
               'Population',
               'Population change from the previous year',
               'Share of persons aged under 15 of the population',
               'Share of persons aged 15 to 64 of the population',
               'Share of persons aged over 64 of the population',
               'Share of Swedish-speakers of the population',
               'Share of foreign citizens of the population',
               'Excess of births',
               'Intermunicipial migration gain/loss',
               'Number of families',
               'Number of household-dwelling units',
               'Share of household-dwelling units living in terraced houses and detached houses',
               'Share of household-dwelling units living in rental dwellings',
               'Share of persons aged 15 or over with at least upper secondary qualifications',
               'Share of persons aged 15 or over with tertiary level qualifications',
               'Employed labour force resident in the area',
               'Employment rate',
               'Share of persons working in their municipality of residence',
               'Proportion of unemployed among the labour force',
               'Proportion of pensioners of the population',
               'Economic dependency ratio',
               'Number of workplaces in the area',
               'Share of workplaces in primary production',
               'Share of workplaces in secondary production',
               'Share of workplaces in services',
               'Workplace self-sufficiency',
               'Annual contribution margin',
               'Loan stock',
               'Group loan stock',
               'Educational and cultural activities',
               'Social and health care activities')

vars <-  c('Deg.Urbanisation',
           'Popul',
           'Popul.Growth',
           'Prop.Under15',
           'Prop.15to64',
           'Prop.Over64',
           'Prop.Swedish',
           'Prop.Foreign',
           'Excess.Births',
           'Migr.Gain',
           'Families',
           'Households',
           'Prop.Households.Terr.Det',
           'Prop.Households.Rental',
           'Prop.Educ.Degree2',
           'Prop.Educ.Degree3',
           'Labour.Force',
           'Empl.Rate',
           'Prop.Empl.Muni',
           'Prop.Unempl',
           'Prop.Pensioners',
           'Depend.Ratio',
           'Jobs.Muni',
           'Prop.Primary.Sector',
           'Prop.Secondary.Sector',
           'Prop.Services.Sector',
           'Jobs.Self.Suff',
           'Contr.Margin',
           'Loan.Stock',
           'Group.Loan.Stock',
           'Educ.Cult.Activity',
           'Soc.Health.Activity')

years <- c(2020,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2019,
           2020,
           2020,
           2020,
           2019,
           2019,
           2019,
           2019,
           2019,
           2020,
           2020,
           2020,
           2020,
           2020)

units <- c("%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "%",
           "%",
           "Persons",
           "Persons",
           "Number",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Ratio",
           "Number",
           "%",
           "%",
           "%",
           "Ratio",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita")

var.df <- data.frame(vars, vars.orig, years, units)

kable(var.df, format = "latex", align = "lll", col.names = c("Variable", "Original Variable","Year","Unit"), caption = "Description of all used variables. Variable is the name of the variable that is used in this report. Original Variable is the original variable given in Tilastokeskus' dataset, Year is the timestamp of the data point, and Unit is the unit in which the variable is measured in.") %>%
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = c("HOLD_position", "scale_down"), position = "left")

```

## Descriptive Statistics

Table \@ref(tab:descstats) shows descriptive statistics for each variable in the data. We can see that many of the variables have quite large variance. This is somewhat expected as the data is a cross-section of 314 Finnish municipalities, so we have very different types of municipalities in the data.

The heterogeneity of the data can be seen from e.g., the population variable, where we have a minimum of 105 inhabitants (Sottunga) and a maximum of 658457 inhabitants (Helsinki). The median population is 6134 and the 3rd quartile is 15145, so it is clear that Finland is a country with mostly small municipalities but a few huge ones.

An interesting observation is also that in some municipalities, only 20.1% work locally inside their own municipality (Kauniainen), while in other the share is as large as 91.7% (Kuusamo). Moreover, in some municipalities, over a fifth is unemployed (21.6% in Ilomantsi) while in other places, the unemployment rate is as low as 3.7% (Föglö).

It is also clear that Finland is a country where in some municipalities, people mostly speak Swedish (92.4% in Sottunga) while in many other, not even a permille speak Swedish (from relatively big municipalities, e.g., Iisalmi). Furthermore, in Sottunga, over a forth (25.7%) are foreign citizens while in Merijärvi, foreign citizens only comprise 0.3% of the population.

```{r descstats, echo=FALSE}
stats <- muni.latest[, -1] %>% apply(2, summary) %>% t() %>% as.data.frame()
stats$Std <- muni.latest[, -1] %>% apply(2, sd)

stats.df <- data.frame(vars, stats)
rownames(stats.df) <- NULL

kable(stats.df, format = "latex", align = "lrrrrrr", caption = "Descriptive Statistics of the variables. Variable is the name of the variable, Min is the minimum, 1st Qu. is the 1st quartile, Median is the median, Mean is the arithmetic mean, 3rd Qu. is the 3rd quartile, Max is the maximum and Std. is the sample standard deviation.", digits = 1, col.names = c("Variable", "Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max", "Std.")) %>% 
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position", position = "left")

```

## Distribution Plots

```{r, include=FALSE}

ages <- muni.latest %>% 
  select(Prop.Under15, Prop.15to64, Prop.Over64)
ages <- melt(ages)

living <- muni.latest %>% 
  select(Prop.Households.Terr.Det, Prop.Households.Rental)
living <- melt(living)

employment <- muni.latest %>% 
  select(Empl.Rate, Prop.Unempl, Prop.Pensioners)
employment <- melt(employment)

sectors <- muni.latest %>% 
  select(Prop.Primary.Sector, Prop.Secondary.Sector, Prop.Services.Sector)
sectors <- melt(sectors)

```

To analyze how variables are distributed, we use Kernel density estimation to estimate the probability densities for several variable groups.

Figure \@ref(fig:agekde) shows the probability density estimates for age groups. Age groups are shares of population aged below 15, 15 to 64, and over 64. We can see that ...

Figure \@ref(fig:housingkde) shows the probability density estimates for housing types. Housing types are shares of population living in either terraced or detached houses and in rental apartments. The Figure revels that ...

Figure \@ref(fig:employmentkde) shows the probability density estimates for employment measures. Employment measures are the employment rate, the proportion of population that are unemployed, and the proportion of population that are pensioners. The distributions show that ...

Figure \@ref(fig:sectorkde) shows the probability density estimates for the weights of economic sectors in the municipalities. Sectors are the primary-, secondary-, and services sectors. We can interpret that ...

```{r agekde, echo=FALSE, fig.cap='Probability density estimates for age groups', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r housingkde, echo=FALSE, fig.cap='Probability density estimates for housing types', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Housing Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r employmentkde, echo=FALSE, fig.cap='Probability density estimates for employment related measures',fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Measures") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

```{r sectorkde, echo=FALSE, fig.cap='Probability density estimates for economic sectors', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(sectors, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

```

# Bivariate Analysis

## Linear Dependencies

To analyze bivariate dependencies in the data, we assessed linear relationships between the variables with Pearson's correlation analysis. Figure \@ref(fig:corrplot) shows a heatmap of the correlations between all variables.

The middle of the heatmap shows that a few of the variables are perfectly positively correlated. Those variables are Families, Popul, Households, and Labour.Force. Jobs.Muni is also nearly perfectly positively correlated with the former variables (correlation coeffcient ranges from 0.98 to 0.99). This observation is not a surprise - we can expect that these variables go hand in hand. Excess.Births is relatively strongly positively correlated with all former variables (correlation coefficient ranges from 0.6 to 0.66). It is believable that excess birth concentrates in bigger municipalities with many other families and job opportunities.

Another cluster of strong positive correlations is in the bottom left corner of the heatmap. Prop.Over64 has correlation of 0.99 with Prop.Pensioners, which is expected, as people tend to start their pension at an age around 65-70. Moreover, Soc.Health.Activity is strongly positively correlated with Prop.Pensioners, Prop.Over64, and Depend.Ratio. This means that the more elderly people live in a municipality, the more the municipality spends on social and health activities.

There are also variables with strong negatively correlations. In the bottom right corner, we see that Prop.15to64 is negatively associated with Soc.Health.Activity, Depend.Ratio, Prop.Pensioners, and Prop.Over64. This means that in contrast to municipalities with old people, municipalities with many 15 to 64 year-olds tend to spend less on social and health activities. The same applies to the variable Prop.Below15.

Interesting observations include the correlations between Prop.Households.Terr.Det and Prop.Primary.Sector, Education (Prop.Educ.Degree2 & Prop.Educ.Degree3) and Deg.Urbanisation, and Prop.Households.Rental and a handful of variables (Jobs.Muni, Labour.Force, Households, Popul, and Families). First, it seems that people tend to live in terraced or detached houses in municipalities that are driven by the primary sector. Second, education appears to be associated with high levels of urbanisation. This is most likely due to the fact that high schools, universities, and high-profile jobs (where highly educated people work) tend to locate in bigger cities that are urbanising the fastest. Lastly, rental housing tends to be linked with jobs, population size, and households. This could be due to the fact that rental housing is more common in big cities, where there tend to be relatively more young people and apartments are more costly.

```{r corrplot, echo=FALSE, fig.dim=c(18,18), fig.cap="Heatmap of Pearson'n correlation between all variables. Red color indicates stron positive correlation while blue color indicates strong negative correlation. White color indicates no correlation.", optipng = '-o7'}
corr <- cor(muni.latest[, -1])
ggcorrplot(corr, hc.order = TRUE, type = "full", lab = TRUE, title = "Heatmap of Correlations", lab_size = 3.5)

```

# Principal Component Analysis (PCA)

## Motivation

PCA is a technique that is used for reducing dimensionality in numeric data while preserving as much as possible of the information (i.e., variation) contained in the original data. As our variables were numeric, and the main objective was to reduce dimensionality in the data, PCA was the best choice. Moreover, we had a prior belief or hypothesis that differences in municipalities could arguably be explained in quite simple terms like center of growth or pensioner municipality, so PCA seemed like an interesting tool to test the hypothesis.

## Scree Plot

Figure \@ref(fig:screeplot) shows what proportion of variance is explained by each PC along with the cumulative proportion of variance explained up to that PC. The scree plot shows that we can explain close to 55% of all variation using only the first two PCs. Furthermore, close to 70% of the variation can be explained by using the first four PCs. In this report, we do further analysis with these four PCs. We have indicated those PCs in the plot with a green color.

```{r screeplot, echo=FALSE, fig.cap="Variance explained by each Princial Component (PC) up to the 20th PC. The bars represent proportion of variance explained and the line with data points indicates the cumulative proportion of variance explained up to that PC. This type of a plot is also known as a Scree Plot.", fig.align='left', optipng = '-o7'}

muni.latest.pca <- muni.latest %>% column_to_rownames(var = "Area")
muni.pca <- prcomp(muni.latest.pca, center = T, scale. = T)

muni.pca.summary <- summary(muni.pca)
scree.df <- muni.pca.summary$importance[, 1:20] %>% t() %>% as.data.frame()
scree.df$PC <- 1:20
scree.df$Included <- c(rep("Yes",4), rep("No",16))
scree.df <- scree.df[, c(4,2,3,5)]
rownames(scree.df) <- NULL
colnames(scree.df) <- c("PC", "Prop.var", "Cum.prop", "Included")

ggplot(scree.df) +
  geom_bar(aes(x = PC, y = Prop.var, fill = Included), stat = "identity") +
  scale_fill_manual(values = c("#440154", "#3dbc74")) +
  geom_line(aes(x = PC, y = Cum.prop)) +
  geom_point(aes(x = PC, y = Cum.prop)) +
  geom_hline(yintercept = 0.70, linetype = 2) +
  annotate("text", x=10.5, y=0.73, label= "70% of variance explained", size = 3) +
  labs(x = "Principal Component", y = "Variance Explained (%)", title = "PCA Scree Plot") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_light()

```

## Principal Components 1 and 2

Figure \@ref(fig:scoreplot1) shows the scores for PCs 1 and 2 for all municipalities that have a top. We can clearly see four clusters of municipalities. The first cluster comprises only of Helsinki - Finland's capital city. The second cluster comprises of Espoo, Vantaa, Tampere, Oulu, and Turku - large cities around Finland. The third cluster comprises of Jomala, Lemland, Luoto, Lumparland, and Saltvik - small Swedish speaking municipalities in rural areas. The forth cluster comprises of Puolanka, Hyrynsalmi, Posio, Rautavaara, and Rääkkylä - small pensioner municipalities in rural areas.

For a deeper characterisation of the revealed clusters, we must look at the loading directions plot, which shows what loadings the different variables have for PCs 1 and 2. Figure \@ref(fig:loadingsplot1) shows the corresponding plot.

We can see that variables that have the most negative loading for PC1 are Prop.15to64, Prop.Educ.Degree3, Families, and Popul while Prop.Pensioners, Prop.Over64, Depend.Ratio, Soc.Health.Activity have strongest positive loadings. This means that PC1 can be interpreted as an index of youthfulness of a municipality - youthful municipalities are highly populated with young, educated working people and families while old municipalities are filled with pensioners.

Considering the PC2 dimension, we see that most negatively loaded variables are Prop.Households.Rental, Prop.Empl.Muni, Prop.Unempl, and Jobs.Self.Suff while Empl.Rate, Prop.Households.Terr.Det, Prop.Under15, and Prop.Swedish have strongest positive loadings. PC2 is more difficult to interpret but a clear sign is that Swedish speaking municipalities load positively for PC2. Living on a rent and unemployment indicate lower economic prosperity than living in a terraced or detached house or employment. Perhaps a high number of children is also a sign of economic wellbeing, since having children can be expensive these days. Therefore, PC2 could be an index of economic prosperity. Big cities have a lot of smaller aparments for rent and many younger people like students live there, which could explain why these cities score quite low in PC2 dimension in comparison to smaller cities.

```{r scoreplot1, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading for any of PC1 and PC2", fig.align='left', optipng = '-o7'}

loads <- muni.pca$x[, 1:2] %>% as.data.frame() %>% rownames_to_column() %>% as_tibble() %>% dplyr::rename(Area = rowname)

greatest5.PC1 <- loads %>% 
  arrange(desc(PC1)) %>% 
  head(5)

smallest5.PC1 <- loads %>% 
  arrange(PC1) %>% 
  head(5)

greatest5.PC2 <- loads %>% 
  arrange(desc(PC2)) %>% 
  head(5)

smallest5.PC2 <- loads %>% 
  arrange(PC2) %>% 
  head(5)

loads.reduced <- full_join(greatest5.PC1, smallest5.PC1) %>% full_join(greatest5.PC2) %>% full_join(smallest5.PC2)

ggplot(loads.reduced, aes(x = PC1, y = PC2, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-25,10)) +
  ylim(c(-20,15)) +
  ggtitle("PCA Scores, 1st and 2nd PC") +
  theme_light()

```

```{r loadingsplot1, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Loading directions for PC1 and PC2", fig.align='left', optipng = '-o7'}

ggplot2::autoplot(muni.pca, 
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.4, 0.4)) +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  labs(x = "Standardized PC1 (33.6% explained var.)",
       y = "Standardized PC1 (20.9% explained var.)",
       title = "PCA Loadings, 1st and 2nd PC") +
  theme_light()

```

Corresponding figures for PC3 and PC4 are shown in the Appendix.

# Discussion and Conclusions

In this report, we have used PCA to reduce dimensionality in demographic data of Finnish Municipalities provided by Tilastokeskus. The results indicate that although Finland has very heterogeneous municipalities, some conclusions can be made after projecting the data in the principal component space. PC1 explains 31.7% of the variance and can be seen as an index of youthfulness of a municipality. PC2 explains 20.5% of the variance and can be interpreted as an index of economic prosperity of a municipality. By plotting the PC1 and PC2 scores for only those municipalities that have a top 5 absolute loading for any of the two components resulted in four clear clusters of municipalities, which demonstates that the objective of the study was accomplished.

Limitations of this study can be seen as the following. First, PC1 and PC2 do not explain a high percentage of the variance (above 80%). Hence, PCA might not be a suitable method to draw conclusions from. Second, using population, employed persons and some other absolute measures might have alienated Helsinki from the other cities and biased the analysis a bit. Perhaps leaving out Helsinki would have been a wise choice. Third, highly correlated variables could have been reduced to only one. In other words, perhaps it would have been wise to remove unnecessary variables. Further research on this topic could address these potential shortcomings or use the findings of the report for more advanced or continued work.

\newpage
# Appendix

Figure \@ref(fig:scoreplot2) shows the scores for PCs 3 and 4.

Figure \@ref(fig:loadingsplot2) shows the loading directions for PCs 3 and 4.

## Principal Components 3 and 4

```{r scoreplot2, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading for any of PC3 and PC4", fig.align='left', optipng = '-o7'}

loads <- muni.pca$x[, 3:4] %>% as.data.frame() %>% rownames_to_column() %>% as_tibble() %>% dplyr::rename(Area = rowname)

greatest5.PC3 <- loads %>% 
  arrange(desc(PC3)) %>% 
  head(5)

smallest5.PC3 <- loads %>% 
  arrange(PC3) %>% 
  head(5)

greatest5.PC4 <- loads %>% 
  arrange(desc(PC4)) %>% 
  head(5)

smallest5.PC4 <- loads %>% 
  arrange(PC4) %>% 
  head(5)

loads.reduced <- full_join(greatest5.PC3, smallest5.PC3) %>% full_join(greatest5.PC4) %>% full_join(smallest5.PC4)

ggplot(loads.reduced, aes(x = PC3, y = PC4, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-5,5)) +
  ylim(c(-8,8)) +
  ggtitle("PCA Scores, 3rd and 4th PC") +
  theme_light()

```

```{r loadingsplot2, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Loading directions for PC3 and PC4", fig.align='left', optipng = '-o7'}

ggplot2::autoplot(muni.pca,
                  x = 3,
                  y = 4,
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.3, 0.3)) +
  scale_y_continuous(limits = c(-0.3, 0.3)) +
  labs(x = "Standardized PC3 (10.2% explained var.)",
       y = "Standardized PC4 (7.2% explained var.)",
       title = "PCA Loadings, 3rd and 4th PC") +
  theme_light()

```

\newpage
# References